Poincaré and SimBio: a versatile and extensible Python ecosystem for modeling systems.

### 2023-12-19 article.qmd
Changed abstract

---
title: "Poincaré and SimBio: a versatile and extensible Python ecosystem for modeling systems."
abstract: >
    Chemical Reaction Networks (CRNs) are a fundamental concept of modelling in numerous fields,
    including systems biology, biochemistry, chemical engineering and epidemiology.
    Formulating them in a higher-level description
    allows to solve them with different approaches,
    such as deterministic or stochastic simulations.
    Existing Python tools for CRN modeling wrap external C/C++ libraries for modelling and simulation,
    which limits their extensibility to new simulation approaches
    or to reuse packages from the vast Python community.

    This work introduces two novel Python packages:
    Poincaré, for definition and simulation of dynamical systems,
    and SimBio, which builds on top of Poincaré to
    facilitate the definition of Chemical Reaction Networks (CRNs),
    and add support for the Systems Biology Markup Language (SBML).

    Poincaré and SimBio are pure Python packages,
    which allows to extend their simulation capabilities by writing Python or leveraging other Python packages.
    But this does not compromise performance,
    as code can be Just-In-Time compiled with Numba,
    and benchmarks show comparable or superior performance to other tools.

    Additionally,
    to ensure a user-friendly experience,
    they use of standard typed modern Python syntax
    that provide a seamless integration with Integrated Development Environments (IDEs),
    enhances code analysis tools, enabling error identification before execution,
    and facilitates code completion and refactoring.

    We believe that this Python-centric approach positions Poincaré and SimBio as valuable tools for the modeling community.

>>>>

---
title: "Poincaré and SimBio: a versatile and extensible Python ecosystem for modeling systems."
abstract: >
    Chemical Reaction Networks (CRNs) play a pivotal role in diverse fields 
    such as systems biology, biochemistry, chemical engineering, and epidemiology. 
    High-level modelling of CRNs enables various simulation approaches, 
    including deterministic and stochastic methods. 
    However, existing Python tools for CRN modelling typically wrap external C/C++ libraries for modelling and simulation, 
    limiting their extensibility and integration with the broader Python ecosystem.
    In response, we developed Poincaré and SimBio, 
    two novel Python packages for the definition and simulation of dynamical systems and CRNs. 
    Poincaré serves as a foundation for dynamical system modelling, 
    while SimBio extends this functionality to CRNs, 
    including support for the Systems Biology Markup Language (SBML). 
    Poincaré and SimBio are developed as pure Python packages 
    enabling users to easily extend their simulation capabilities 
    by writing new or leveraging other Python packages.
    Moreover, this does not compromise the performance, 
    as code can be Just-In-Time compiled with Numba. 
    Our benchmark tests using curated models from the BioModels repository, 
    demonstrate that these tools may provide a potentially superior performance advantage, 
    compared to other existing tools.

    Additionally, 
    to ensure a user-friendly experience,
    our packages use standard typed modern Python syntax
    that provides a seamless integration with Integrated Development Environments (IDEs).

    Python-centric approach significantly enhances code analysis, 
    error detection, and refactoring capabilities, 
    positioning Poincaré and SimBio as valuable tools for the modelling community.
